---
title: "Bayes's Rule and Sampling"
author: "서용덕"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("Rlab")
```

## Bernoulli distribution model

The sample $Y$ is obtained from a Bernoulli distribution
$$
Y \sim Bern(\theta)
$$
where
$$
  P(Y|\theta) = Bern(Y|\theta) = \left\{ 
    \begin{array}{c}
      \theta & \mbox{if }  Y=1\\
      1-\theta & \mbox{if } Y = 0
    \end{array}
  \right.
$$
or equivalently
$$
  P(Y|\theta) = \theta^Y (1-\theta)^{1-Y}
$$


In R, we can get random samples from Bern(0.5):
```{r}
  theta = 0.5
  nsamples = 11
  y = rbern(nsamples, theta)
  print(y)
  
```
```{r}
  theta = 0.7
  nsamples = 1001
  y = rbern(nsamples, theta)
  hist(y, main="histogram of Bernoulli samples")
  
```

```{r}
  theta = 0.5
  nsamples = 1001
  y = rbern(nsamples, theta)
  hist(y)
  
```

```{r}
  theta = 0.3
  nsamples = 1001
  y = rbern(nsamples, theta)
  hist(y)
  
```

```{r}
  theta = 0.1
  nsamples = 1001
  y = rbern(nsamples, theta)
  hist(y)
  
```

As can be seen, the random outputs depend on the value of the parameter $\theta$


## Posterior probability
```{r}
  data = c(1,1,1,1,0)  # random outputs
```

Q: Given this data, what is the posterior probability of $\theta$, the probability of producing "1" in the experiment?

```{r}
  vtheta <- c(.3, .4, .5, .6, .7) # possible values of theta values
  p_theta <- c(.3, .7, 1, .7, .3)
  p_theta = p_theta / sum(p_theta)
  barplot(p_theta, names=vtheta)
```

```{r}
likelihood = vtheta * vtheta * vtheta * (1- vtheta) * (1-vtheta)
Punnormalized = likelihood * p_theta
Z = sum(Punnormalized)
posterior = Punnormalized / Z

barplot(posterior, names=vtheta)
```

```{r}
ratio = posterior / p_theta
print(ratio)
barplot(ratio, names=vtheta, col="#700000")
abline(h=1)
```
```{r}
barplot(rbind(posterior, p_theta),
        beside=T,
        col=c("#700000", "#000070"),
        names=vtheta)
```



## Beta distribution for $\theta \in [0, 1]$

$$
  \theta \sim Beta(\alpha, \beta)
$$
```{r}
  t = seq(0, 1, length=101)
  alpha = 4
  beta  = 2
  pbeta = dbeta(t, alpha, beta)
  plot(t, pbeta, col="#00007f")
```


```{r}
# Define the range
p <- seq(0, 1, length = 100)

# Plot the first Beta distribution
plot(p, dbeta(p, 2, 5), type = 'l', col = 'blue', 
     main = 'Beta Distributions', 
     xlab = 'x', 
     ylab = 'Density')

# Add more Beta distributions
lines(p, dbeta(p, 5, 2), col = 'red')
lines(p, dbeta(p, 2, 2), col = 'green')

# Add a legend
legend('topright', legend = c('Beta(2, 5)', 'Beta(5, 2)', 'Beta(2, 2)'), 
       col = c('blue', 'red', 'green'), lty = 1)
```

```{r}
# Define the range
p <- seq(0, 1, length = 50)

# Plot the first Beta distribution
plot(p, dbeta(p, 12, 25), type = 'l', col = 'blue', 
     main = 'Beta Distributions', 
     xlab = 'x', 
     ylab = 'Density')

# Add more Beta distributions
lines(p, dbeta(p, 25, 12), col = 'red')
lines(p, dbeta(p, 20, 20), col = 'green')

# Add a legend
legend('topright', legend = c('Beta(12, 25)', 'Beta(25, 12)', 'Beta(20,20)'), 
       col = c('blue', 'red', 'green'), lty = 1)
```



## Posterior with Beta grid
The same data
```{r}
  data = c(1,1,1,0,0)  # random outputs
```

Q: Given this data, what is the posterior probability of $\theta$, the probability of producing "1" in the experiment?

```{r}
  vtheta <- seq(0, 1, length = 21) # possible values of theta values
  p_theta <- dbeta(vtheta, 5, 5)   # probability density is obtained
  p_theta = p_theta / sum(p_theta) # modify it to use as a PMS
  barplot(p_theta, names=round(vtheta, digits=2), main="Prior distribution of theta")
```

```{r}
# likelihood = vtheta * vtheta * vtheta * (1- vtheta) * (1-vtheta)
likelihood = 1
for (i in 1:length(data)) {
  if (data[i] == 1) {
    likelihood = likelihood * vtheta
  }
  else{ 
    likelihood = likelihood * (1 - vtheta)
  }
}

barplot(likelihood, names=round(vtheta, 2), main="Likelihood function")
```

Notice that the sum of the likelhood is not 1:
```{r}
print(paste("sum of likelihood = ", sum(likelihood)))
```

<<<<<<< HEAD
Now, we compute the posterior
=======
Sanity check - the sum must be 1 for every i.
>>>>>>> 3b8a909d1c3ad14d1aa128b827105bbe757b16b0
```{r}
Punnormalized = likelihood * p_theta
Z = sum(Punnormalized)
posterior = Punnormalized / Z
print(paste("Z = ", Z))
```
```{r}
barplot(posterior, names=round(vtheta, digits=2), main="Posterior")
```

```{r}
ratio = posterior / p_theta
barplot(ratio, names=vtheta, col="#700000")
abline(h=1)
```
```{r}
barplot(rbind(posterior, p_theta),
        beside=T,
        col=c("#700000", "#0000f0"),
        names=vtheta)
```
## MLE, MAP, Posterior distribution

* MLE: Maximum likelihood estimate. The theta value where the likelihood function is maximum.
* MAP: Maximum location of the posterior distribution

```{r}
index_mle = which.max(likelihood)
mle = vtheta[index_mle]
print(paste("MLE = ", mle))
```

```{r}
barplot(likelihood, names=round(vtheta,2),
        main=paste("Likelihood; max @ theta = ", mle)
        )
```
```{r}
index_map = which.max(posterior)
MAP = vtheta[index_map]
print(paste("MAP=", MAP))

index_map2 = which.max(Punnormalized)
MAP2 = vtheta[index_map]
print(paste("MAP=", MAP2))
```


```{r}
barplot(posterior, names=round(vtheta,2),
        main=paste("Posterior; max @ theta = ", round(MAP,2))
)
```


# Symbolic computation

bernoulli likelihood * beta prior = beta posterior

$$
  Beta(\theta|a, b) = \frac{\theta^{a-1} (1-\theta)^{b-1} }{B(a,b)}
$$
where $B(a,b)$ is a normalization constant depending on $a$ and $b$ defined by:
$$
  B(a,b) = \int_0^1 \theta^{a-1} (1-\theta)^{b-1} d\theta
$$
or
$$
  B(a,b) = \frac{\Gamma(a)\Gamma(b)}{\Gamma(a+b)} \quad\quad
  \Gamma(a) = \int_0^\infty t^{a-1} e^{-t} dt
$$




























